---
title: "cross_validation"
author: "Catherine"
date: "11/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Although hypothesis tests provide a way to compare nested linear models, in many situations the approaches under consideration don’t fit nicely in this paradigm. Indeed, for many modern tools and in many applications, the emphasis lies on prediction accuracy rather than on statistical significance. In these cases, cross validation provides a way to compare the predictive performance of competing methods.


```{r}
library(tidyverse)
library(modelr)
library(mgcv)
```

```{r}
nonlin_df = tibble(
  id = 1:100,
  x = runif(100, 0, 1),
  y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
)

ggplot(nonlin_df, aes(x = x, y = y)) + geom_point() + theme_bw()

train_df = sample_n(nonlin_df, 80)
test_df = anti_join(nonlin_df, train_df, by = "id")

ggplot(train_df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_point(data = test_df, color = "red")
```
```{r}
lin_mod = lm(y ~ x, data = train_df)
nonlin_mod = mgcv::gam(y ~ s(x), data = train_df)
wiggly_mod = mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)

train_df %>% 
  add_predictions(nonlin_mod) %>% 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")

train_df %>% 
  add_predictions(wiggly_mod) %>% 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

```{r}
train_df %>% 
  gather_predictions(lin_mod, nonlin_mod, wiggly_mod) %>% # add predictions all in one dataframe
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_wrap(~model)
```

```{r}
# compute model quality for a given dataset - `root-mean-squared-error` (lower the better)
rmse(lin_mod, test_df)
rmse(nonlin_mod, test_df)
rmse(wiggly_mod, test_df)
```

Iterate!

```{r}
cv_df = 
  crossv_mc(nonlin_df, 100) # 100 version for training and testing dataset
cv_df %>% pull(train) %>% .[[1]] %>% as_tibble()

#  fit my candidate models above and assess prediction accuracy as I did for the single training / testing split
cv_df =
  cv_df %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))

cv_df = 
  cv_df %>% 
  mutate(lin_mod    = map(train, ~lm(y ~ x, data = .x)),
         nonlin_mod = map(train, ~mgcv::gam(y ~ s(x), data = .x)),
         wiggly_mod = map(train, ~gam(y ~ s(x, k = 30), sp = 10e-6, data = .x))) %>% 
  mutate(rmse_lin    = map2_dbl(lin_mod, test, ~rmse(model = .x, data = .y)), 
         rmse_nonlin = map2_dbl(nonlin_mod, test, ~rmse(model = .x, data = .y)),
         rmse_wiggly = map2_dbl(wiggly_mod, test, ~rmse(model = .x, data = .y)))
# map2_dbl you get the number with`_dbl`
```

summerize results

```{r}
cv_df %>% 
  select(.id, starts_with("rmse")) %>% 
  gather(key = model, value = rmse, rmse_lin:rmse_wiggly) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
  
```

Child grwoth example

```{r}
child_growth = read_csv("./data/nepalese_children.csv")
```

```{r}
child_growth %>% 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .5)
```

Is it linear?? Let's prove

```{r}
child_growth =
  child_growth %>% 
  mutate(weight_sp = (weight > 7) * (weight - 7))
```
```{r}
lin_mod = lm(armc ~ weight, data = child_growth)
pwl_mod = lm(armc ~ weight + weight_sp, data = child_growth)
nonlin_mod = gam(armc ~ s(weight), data = child_growth)
```

plot models

```{r}
child_growth %>% 
  gather_predictions(lin_mod, pwl_mod, nonlin_mod) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(~model)
```

```{r}
cv_df =
  crossv_mc(child_growth, 100) %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))
```

```{r}
cv_df = 
  cv_df %>% 
  mutate(lin_mod = map(train, ~lm(armc ~ weight, data = .x)),
         pwl_mod = map(train, ~lm(armc ~ weight + weight_sp, data = .x)),
         nonlin_mod = map(train, ~gam(armc ~ s(weight), data = as_tibble(.x)))) %>% 
  mutate(rmse_lin    = map2_dbl(lin_mod, test, ~rmse(model = .x, data = .y)),
         rmse_pwl = map2_dbl(pwl_mod, test, ~rmse(model = .x, data = .y)),
         rmse_nonlin = map2_dbl(nonlin_mod, test, ~rmse(model = .x, data = .y)))
```

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```



